Recurrence spike spectra

### 2022-07-19 Manuscript/inter_spike_manuscript.tex
Prepare for entropy journal

Some of these recurrence quantification measures, the entropy of diagonal lines and the entropy of 
recurrence times, can be related to the Kolmogorov-Sinai entropy \cite{march2005,baptista2010}. However, these quantifiers have a free parameter, the minimal considered line length, and 
are usually biased, due to the finite size of the RP and thickened diagonal lines, which needs to be corrected \cite{Kraemer2019}. Moreover, the mentioned statistics cannot account for 
changing regular (non-chaotic) dynamics, such as period-doubling bifurcations.
A rather simple idea is to look at the $\tau$-recurrence rate of the RP ($\tau$-RR, Eq.~\ref{eq_tau_rr}) \cite{marwan2002pla,Zbilut2008}.
This is the density of recurrence points along the diagonals of the recurrence matrix, as a function of the distance $\tau$ (sampling units) to the main diagonal:
\begin{equation}\label{eq_tau_rr}
\tau\text{-RR}(\varepsilon) = RR(\tau, \varepsilon) = \frac{1}{N-\tau} \sum_{i=1}^{N-\tau	} R_{i,i+\tau}.
\end{equation}
$\tau$-RR serves as an estimator for the probability that the system recurs after time $\tau \Delta t$, with $\Delta t$ being the sampling time of the trajectory 
$\vec{x}_i = \vec{x}(\Delta t \cdot i),\, i=1,\ldots,N$. 
\cite{Zbilut2008} pointed out that $\tau$-RR could be used as a plugin value for the auto-correlation function $C(\tau)$ and, hence, via the Wiener-Khinchim theorem a 
``generalized'' powerspectrum can be obtained. This is reasonable, since the average distances for a given lag $\tau$ 
\begin{equation}
\overline{D}(\tau) = \frac{1}{N-\tau}\sum_{i=1}^{N-\tau} D_{i, i+\tau}
\end{equation}
can be directly read from the distance matrix $\mathbf{D}$ and is also preserved in its thresholded version $\tau$-RR. There are clear advantages for a recurrence-derived 
powerspectrum, i.e., Fourier transforming (FT) $\tau$-RR (Fig.~\ref{fig_tau_rr_spectrum_example}D), instead 
of $C(\tau)$ (Fig.~\ref{fig_tau_rr_spectrum_example}C): There are no assumptions for stationarity or sampling, when constructing a RP.
Furthermore, the correlation structures of higher dimensional spaces can be resolved in the recurrence-derived Fourier-spectrum.\\

>>>>

Some of these recurrence quantification measures, the entropy of diagonal lines and the entropy of 
recurrence times, can be related to the Kolmogorov-Sinai entropy \cite{march2005,baptista2010}. However, these quantifiers have a free parameter, the minimal considered line length, and 
are usually biased, due to the finite size of the RP and thickened diagonal lines, which needs to be corrected \cite{Kraemer2019}. Moreover, the mentioned statistics cannot account for 
changing regular (non-chaotic) dynamics, such as period-doubling bifurcations.
A rather simple idea is to look at the $\tau$-recurrence rate of the RP ($\tau$-RR, Eq.~\ref{eq_tau_rr}) \cite{marwan2002pla,Zbilut2008}.
This is the density of recurrence points along the diagonals of the recurrence matrix, as a function of the distance $\tau$ (sampling units) to the main diagonal:
\begin{equation}\label{eq_tau_rr}
\tau\text{-RR}(\varepsilon) = RR(\tau, \varepsilon) = \frac{1}{N-\tau} \sum_{i=1}^{N-\tau	} R_{i,i+\tau}.
\end{equation}
$\tau$-RR serves as an estimator for the probability that the system recurs after time $\tau \Delta t$, with $\Delta t$ being the sampling time of the trajectory 
$\vec{x}_i = \vec{x}(\Delta t \cdot i),\, i=1,\ldots,N$. 
\citet{Zbilut2008} pointed out that $\tau$-RR could be used as a plugin value for the auto-correlation function $C(\tau)$ and, hence, via the Wiener-Khinchim theorem a 
``generalized'' powerspectrum can be obtained. This is reasonable, since the average distances for a given lag $\tau$ 
\begin{equation}
\overline{D}(\tau) = \frac{1}{N-\tau}\sum_{i=1}^{N-\tau} D_{i, i+\tau}
\end{equation}
can be directly read from the distance matrix $\mathbf{D}$ and is also preserved in its thresholded version $\tau$-RR. There are clear advantages for a recurrence-derived 
powerspectrum, i.e., Fourier transforming (FT) $\tau$-RR (Fig.~\ref{fig_tau_rr_spectrum_example}D), instead 
of $C(\tau)$ (Fig.~\ref{fig_tau_rr_spectrum_example}C): There are no assumptions for stationarity or sampling, when constructing a RP.
Furthermore, the correlation structures of higher dimensional spaces can be resolved in the recurrence-derived Fourier-spectrum.\\
